Estimating Neighborhood Rents with Scraped Data

### 2018-07-27 manuscript/manuscript.Rmd
Editing the Helena parts.

We start by briefly reviewing some background relevant to studying neighborhood rental housing. I then outline a process for automated web scraping of a rental listing source like Craigslist using a program called Helena, and compare aggregates from the collected rental listings with other data to assess rent agreement in measures. After describing the data and process, We present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data. We conclude by discussing the promise of web data for subfields like housing studies and the value of R Markdown and GitHub for social science research.

@@

[Helena](http://helena-lang.org) is a high-level programming language aimed at helping end users automate web scraping as a method of data collection. Users write programs in an extension for Google Chrome by demonstrating the first row of data they want to collect using clicks, form interactions and highlighting values. The extension translates an initial demonstration into a starting script written in Helena, which a user can further modify or set to run on a regular schedule. Though Helena programs collect data roughly at the speed they are recorded at, users can run multiple programs in parallel to increase the volume of data collected. Data accumulate on a remote server (soon-to-be AWS) where users can later query either the full table of scraped data or a subset scraped within a specified number of hours.

Helena learns where features of are through the user interface (UI) rather than the underlying document object model (DOM). Though the HTML code of a site may regularly change over the course of regular data scraping, websites less frequently overhaul what their users actually interact with. Helena also uses skip blocks to avoid re-scraping the an already-collected item. These two aspects of Helena provide its programs robustness to common headaches encountered with web scraping by default. Overall, Helena's programming-by-demonstration method reduces time needed building a scraper from scratch, giving social scientists the capacity to quickly scrape a single table of data or build ongoing data processes.

For the UW team's data collection, a Helena program runs 24-hours a day and scrapes listings from the Seattle Craigslist. It is important to note that web scraping does violate Craigslist's Terms of Service, and that we are using all scraped data exclusively for academic research. For a more thorough discussion of this issue, consult page 3 of [Boeing and Waddell (2016)](https://arxiv.org/ftp/arxiv/papers/1605/1605.05397.pdf).

>>>>

We start by briefly reviewing some background relevant to studying neighborhood rental housing. I then outline a process for automated web scraping of a rental listing source like Craigslist using a Programming by Demonstration (PBD) tool called Helena and compare aggregates from the collected rental listings with other data to assess rent agreement in measures. After describing the data and process, We present a set of models predicting neighborhood rent levels to understand different how different temporal and spatial structures fit the data. We conclude by discussing the promise of web data for subfields like housing studies and the value of R Markdown and GitHub for social science research.

@@

[Helena](http://helena-lang.org) is a high-level programming language aimed at helping end users automate web scraping as a method of data collection. Users write programs in an extension for Google Chrome.  To write a new program, they record themselves collecting the first row of the data they want to collect; they use standard browser interactions -- clicking, typing, and interacting with webpage widgets, and they highlight content they want to scrape.  The Helena Chrome extension translates this input demonstration into a scraping program written in the Helena language, which a user can run, edit, or schedule for later runs. Although Helena programs collect data at roughly the speed of the user demonstration, users can run Helena programs in parallel to increase the rate of data collection. Collected data is stored on a remote server, from which users can later download either the full table of scraped data or a subset scraped within a specified window of time.

Helena learns how to find data on a webpage using features of the user-facing webpage interface rather than relying on the stability of the underlying document object model (DOM) structure. Thus, even if the HTML code of a site changes frequently over the course of longitudinal data scraping, as long as the website limits overhauls of the features users see, Helena can often adapt to webpage redesigns. Helena also uses a language construct called the skip block to avoid re-scraping already-collected items. These features make Helena programs robust to issues that commonly plague traditional web scraping workflows. Overall, Helena's demonstration-based method reduces the time needed to build a scraper from scratch, giving social scientists the capacity to quickly scrape a single dataset or to build ongoing data collection processes.

For the UW team's data collection, a Helena program runs 24 hours a day to scrape rental listings from the Seattle Craigslist. It is important to note that web scraping does violate Craigslist's Terms of Service and that we are using all scraped data exclusively for academic research. For a more thorough discussion of this issue, consult page 3 of [Boeing and Waddell (2016)](https://arxiv.org/ftp/arxiv/papers/1605/1605.05397.pdf).
